Trajectories

### 2022-12-01 index.qmd
Replace colour palettes with achromatic-safe colours

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme

options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.ordinal.fill = ggsci::pal_d3()(4),
    ggplot2.ordinal.colour = ggsci::pal_d3()(4),
    ggplot2.discrete.fill = ggsci::pal_d3()(4),
    ggplot2.discrete.colour = ggsci::pal_d3()(4),
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)
```

@@

```{r}
#| label: fig-participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 10 
#| fig-cap: "Date at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1, colour = pal_d3()(1)[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 20, hjust = 1),
        axis.title.x = element_blank(),
    ) +

    participants %>% 
    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
    group_by(time_stamp, age) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(age) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = age)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
    scale_y_continuous(breaks = seq(0, 500, 100)) +
    scale_color_d3() +

@@

```{r}
#| label: fig-participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: fig-participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    pivot_longer(c(doe_catalan, doe_spanish),
                 names_to = "language", values_to = "doe", 
                 names_transform = ~str_to_sentence(str_remove(., "doe_"))) %>% 
    mutate(doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
           id = paste0(id, " (", time, ")")
    ) %>% 
    count(language, doe) %>% 
    ggplot(aes(doe, y = n, label = n, fill = language)) +
    geom_col(colour = "white", size = 1) +
    geom_text(position = position_nudge(y = 5)) +
    facet_wrap(~language, ncol = 1) + 
    labs(
        x = "Degree of exposure to Catalan",
        y = "Number of participants",
        colour = "Language",
        fill = "Language"
    ) +
    theme(
        legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank()
    )
```

```{r}
#| label: fig-participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 15
#| fig-height: 8
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    mutate(
        dominance = ifelse(doe_catalan >= doe_spanish,
                           "Dominant in Catalan",
                           "Dominant in Spanish"),
        id = paste0(id, " (", time, ")"),
        doe_other = 1 - (doe_spanish + doe_catalan)
    ) %>% 
    mutate(rank = rank(doe_catalan)) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe", 
        names_transform = ~str_to_sentence(str_remove(., "doe_"))
    ) %>% 
    mutate(language = factor(language, levels = c("Catalan", "Spanish", "Other"))) %>%
    ggplot(aes(reorder(id, desc(rank)), doe, colour = language, fill = language)) +
    facet_wrap(~dominance, ncol = 1, scales = "free_x") +
    geom_line(aes(group = id), colour = "grey") +
    geom_point(size = 0.5) +
    labs(x = "Participant", y = "Degree of exposure (DoE)", 
         fill = "Language", colour = "Language") +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank()
    )
```
### SES/parental education

@@

```{r}
#| label: tbl-items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(pool, te, item, language, class, semantic_category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq_zipf, n_phon, lv), 
        list(
            n = ~sum(!is.na(.)),
            mean = ~mean(., na.rm = TRUE),
            sd = ~sd(., na.rm = TRUE),
            min = ~min(., na.rm = TRUE),
            max = ~max(., na.rm = TRUE)
        )
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_zipf_n) %>% 
    gt(groupname_col = "language", rowname_col = "list") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd")) %>% 
    fmt_percent(matches("lv_")) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_zipf_mean = md("Mean (SD)"),
        freq_zipf_min = md("Range"),
        n_phon_mean = md("Mean (SD)"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean (SD)"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    ) %>% 
    summary_rows(
        fns = list("Grand mean" = ~mean(.)),
        columns = ends_with("_mean")
    ) %>% 
    summary_rows(
        fns = list("N" = ~sum(.)),
        columns = n,
        formatter = fmt_integer
    )
```

@@

```{r}
#| label: fig-items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(
        fill = ggsci::pal_d3()(4)[1],
        colour = "white"
    ) +
    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
              position = position_nudge(y = 5)) +
    labs(x = "Levenshtein distance", y = "Probability density") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(freq_zipf, lv, language) %>% 
    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(size = 2, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of phonemes and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(n_phon, lv, language) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(shape = 1, stroke = 1) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

We conducted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension). In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes). This is similar to the control conducted by ANCOVAs. Thus, independent variable effects account for collective differences in word learning (unexplained by participants’ or items’ individual differences).

@@

We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO), a relative fit index, via the `loo` R package. LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.

@@

#### Prior predictions

@@

##### Prior-predicted mean

@@

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    scale_fill_brewer(palette = "Reds", labels = c("95%", "89%", "67%", "50%")) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

##### Prior-predictions

@@

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

```{r}
#| lab: fig-responses-age-phonemes
#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
df %>% 
    mutate(
        age = floor(age),
        n_phon = cut_interval(n_phon, n = 4)
    ) %>%
    count(age, n_phon, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~n_phon) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_d3() +
    scale_y_continuous(labels = percent) +
    theme_custom() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| lab: fig-responses-age-exposure
#| fig-cap: "Proportion of responses to each category by age and exposure."
df %>% 
    mutate(
        age = floor(age),
        exposure = cut_interval(exposure, n = 4)
    ) %>%
    count(age, exposure, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~exposure) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_d3() +
    scale_y_continuous(labels = percent) +
    theme_custom() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: tbl-results-loos
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    gt() %>% 
    tab_header(
        title = "",
        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_3`)")
    ) %>% 
    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
    fmt_number(2:9) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("*SE*"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_title(groups = "subtitle")
    )
```

@@

marg_epreds <- predictions(
    model_fit_4,
    newdata = nd,
    ndraws = 25, 
    re_formula = NA
)

>>>>

# set ggplot theme and colour palette
theme_set(theme_custom()) # set custom ggplot theme
clrs <- c("#003f5c", "#58508d", "#bc5090", "#ff6361", "#ffa600")

options(
    knitr.kable.NA = "-",
    knitr.duplicate.label = "allow",
    ggplot2.ordinal.fill = clrs,
    ggplot2.ordinal.colour = clrs,
    ggplot2.discrete.fill = clrs,
    ggplot2.discrete.colour = clrs,
    ggplot2.continuous.fill = ggplot2::scale_color_gradient,
    ggplot2.continuous.colour = ggplot2::scale_color_gradient
)
```

::: {.callout-tip}
Data and materials are available at [OSF](https://osf.io/hy984/), and code is available at [GitHub](https://github.com/gongcastro/trajectories).
:::

@@

```{r}
#| label: fig-participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 10 
#| fig-cap: "Date at which responses were collected"
participants %>% 
    group_by(time_stamp) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    annotate(geom = "text", y = 350, x = ymd("2020-03-15"), label = "COVID-19 lockdown in Spain",
             angle = 90, vjust = 1.5, hjust = 1) +
    geom_line(size = 1, colour = clrs[1]) +
    labs(x = "Date", y = "Number of responses", colour = "Dominant language") +
    scale_y_continuous(breaks = seq(0, 500, 100), limits = c(0, 500)) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_text(angle = 20, hjust = 1),
        axis.title.x = element_blank(),
    ) +

    participants %>% 
    mutate(age = cut(floor(age), breaks = seq(7, 35, 5))) %>% 
    group_by(time_stamp, age) %>% 
    summarise(n = n(), .groups = "drop") %>% 
    group_by(age) %>% 
    mutate(n = cumsum(n)) %>% 
    ggplot(aes(time_stamp, n, colour = age)) +
    geom_vline(xintercept = ymd("2020-03-15"), size = 0.5) +
    geom_line(size = 1) +
    labs(x = "Date", y = "Number of responses", colour = "Age (months(") +
    scale_y_continuous(breaks = seq(0, 500, 50)) +
    scale_colour_manual(values = clrs) +

@@

```{r}
#| label: fig-participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = clrs[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: fig-participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    pivot_longer(
        c(doe_catalan, doe_spanish),
        names_to = "language", values_to = "doe", 
        names_transform = ~str_to_sentence(str_remove(., "doe_"))
    ) %>% 
    mutate(
        doe = cut(doe, breaks = seq(0, 1, 0.1), include.lowest = TRUE),
        id = paste0(id, " (", time, ")")
    ) %>% 
    count(language, doe) %>% 
    ggplot(aes(doe, y = n, label = n, fill = language)) +
    geom_col(colour = "white", size = 1) +
    geom_text(position = position_nudge(y = 5)) +
    facet_wrap(~language, ncol = 1) + 
    labs(
        x = "Degree of exposure to Catalan",
        y = "Number of participants",
        colour = "Language",
        fill = "Language"
    ) +
    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
    theme(
        legend.position = "top",
        panel.grid.major.x = element_blank(),
        panel.grid.minor = element_blank()
    )
```

### SES/parental education

@@

```{r}
#| label: tbl-items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(bvqdev::pool, te, item, language, class, semantic_category))

items %>%
    unnest(list) %>% 
    group_by(language, list) %>% 
    summarise_at(
        vars(freq_zipf, n_phon, lv), 
        list(
            n = ~sum(!is.na(.)),
            mean = ~mean(., na.rm = TRUE),
            sd = ~sd(., na.rm = TRUE),
            min = ~min(., na.rm = TRUE),
            max = ~max(., na.rm = TRUE)
        )
    ) %>% 
    select(-c(n_phon_n, lv_n)) %>% 
    rename(n = freq_zipf_n) %>% 
    gt(groupname_col = "language", rowname_col = "list") %>% 
    fmt_number(matches("freq_|n_phon_mean|n_phon_sd"), decimals = 1) %>% 
    fmt_percent(matches("lv_"), decimals = 1) %>% 
    tab_spanner("Frequency (Zipf)", columns = starts_with("freq_")) %>% 
    tab_spanner("Length (Phonemes)", columns = starts_with("n_phon_")) %>% 
    tab_spanner("Phon. similarity (Levenshtein)", columns = starts_with("lv_")) %>% 
    cols_merge_uncert(col_val = freq_zipf_mean, col_uncert = freq_zipf_sd) %>% 
    cols_merge_uncert(col_val = n_phon_mean, col_uncert = n_phon_sd) %>% 
    cols_merge_uncert(col_val = lv_mean, col_uncert = lv_sd) %>% 
    cols_merge_range(col_begin = freq_zipf_min, col_end = freq_zipf_max) %>% 
    cols_merge_range(col_begin = n_phon_min, col_end = n_phon_max) %>% 
    cols_merge_range(col_begin = lv_min, col_end = lv_max) %>% 
    cols_label(
        list = "",
        n = md("*N*"),
        freq_zipf_mean = md("Mean ± SD"),
        freq_zipf_min = md("Range"),
        n_phon_mean = md("Mean ± SD"),
        n_phon_min = md("Range"),
        lv_mean = md("Mean ± SD"),
        lv_min = md("Range")
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(everything())
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(everything())
    ) %>% 
    summary_rows(
        fns = list("Grand mean" = ~mean(.)),
        columns = ends_with("_mean")
    ) %>% 
    summary_rows(
        fns = list("N" = ~sum(.)),
        columns = n,
        formatter = fmt_integer
    )
```

@@

```{r}
#| label: fig-items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 9
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv) %>% 
    mutate(prop = n/sum(n)) %>% 
    ggplot(aes(x = lv, n)) +
    geom_col(fill = clrs[1], colour = "white") +
    geom_text(aes(label = paste0(n, " (", percent(prop, accuracy = 0.1), ")")),
              position = position_nudge(y = 3), size = 3.5) +
    labs(x = "Levenshtein distance", y = "Probability density") +
    theme(
        panel.grid.major.x = element_blank(),
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    left_join(select(bvqdev::pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(freq_zipf, lv, language) %>% 
    ggplot(aes(freq_zipf, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(size = 2, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of phonemes and phonological similarity"
items %>% 
    left_join(select(bvqdev::pool, item, language)) %>% 
    drop_na(freq_zipf) %>% 
    count(n_phon, lv, language) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language, size = n)) +
    facet_wrap(~language) +
    geom_point(shape = 1, stroke = 1) +
    geom_smooth(method = "lm", formula = "y ~ x", size = 1) +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

We extracted four variables of interest for each observation (describe in previous sections):

* Word lexical frequency ($Frequency$)
* Word length ($Phonemes$)
* Phonological similarity ($Levenshtein$)
* Participant age ($Age$)
* Word exposure index ($Exposure$)

We fitted a series of nested Bayesian ordinal multilevel models to sequentially test the effects of the independent variables (number of phonemes, exposure, age, and Levenshtein) over the three levels of word learning (no, production, comprehension). In all these models participants’ and items’ individual differences were controlled by estimating different parameters for each one (different intercepts and/or slopes). This is similar to the control conducted by ANCOVAs. Thus, independent variable effects account for collective differences in word learning (unexplained by participants’ or items’ individual differences).

@@

We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (*LOO*), a relative fit index, via the `loo` R package. LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_4$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model.

@@

@@

#### Prior-predicted mean

@@

m %>% 
    ggplot(aes(x = age_std, y = .epred)) +
    facet_wrap(~.category) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(
        values = rev(clrs), 
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank()
    )
```

#### Prior-predictions

@@

m %>% 
    count(age_std, .prediction) %>% 
    pivot_wider(names_from = .prediction, values_from = n) %>% 
    mutate(Understands = Understands + `Understands and Says`) %>% 
    pivot_longer(c(No, Understands, `Understands and Says`), names_to = ".prediction", values_to = "n") %>%
    ggplot(aes(x = age_std, y = n, fill = .prediction)) +
    geom_col(position = position_fill()) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition",
        colour = "Category", 
        fill = "Category"
    ) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4)
    ) +
    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        axis.ticks.x = element_line(colour = "black"),
        panel.grid = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

```{r}
#| lab: fig-responses-age-phonemes
#| fig-cap: "Proportion of responses to each category by age and number of phonemes"
df %>% 
    mutate(
        age = floor(age),
        n_phon = cut_interval(n_phon, n = 4)
    ) %>%
    count(age, n_phon, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~n_phon) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
    scale_y_continuous(labels = percent) +
    theme_custom() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| lab: fig-responses-age-exposure
#| fig-cap: "Proportion of responses to each category by age and exposure."
df %>% 
    mutate(
        age = floor(age),
        exposure = cut_interval(exposure, n = 4)
    ) %>%
    count(age, exposure, response) %>% 
    ggplot(aes(age, n, fill = response)) +
    facet_wrap(~exposure) +
    geom_col(position = position_fill()) +
    labs(x = "Age (months)", y = "Proportion of responses", fill = "Response") +
    scale_fill_manual(values = clrs[c(1, 3, 5)]) +
    scale_y_continuous(labels = percent) +
    theme_custom() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: tbl-results-loos
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV). This procedure involves estimating the likelihood of each observation given a model fitted without such observation. The resulting likelihoods of each model were summed to compute the expected log-predicted density (*ELPD*) of the model, and correcting it for the number of parameters in the model to penalised for its the complexity. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*LOO-IC*) of each model along their associated standard errors in paratheses. The first column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_4`)."
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("diff"),
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic")
    ) %>%
    gt() %>% 
    fmt_number(2:9, decimals = 1) %>% 
    cols_merge_uncert(col_val = elpd_diff, col_uncert = se_diff) %>% 
    cols_merge_uncert(col_val = elpd_loo, col_uncert = se_elpd_loo) %>% 
    cols_merge_uncert(col_val = p_loo, col_uncert = se_p_loo) %>% 
    cols_merge_uncert(col_val = looic, col_uncert = se_looic) %>% 
    cols_label(
        model = "Model",
        elpd_loo = md("*ELPD*"),
        p_loo = md("*p*"),
        looic = md("*LOO-IC*"),
        elpd_diff = md("Difference")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels()
    ) %>% 
    tab_style(
        cell_fill(color = "grey", alpha = 0.5),
        cells_body(matches("diff"))
    )
```

@@

marg_epreds <- predictions(
    model_fit_4,
    newdata = nd,
    ndraws = NULL, 
    re_formula = NA
)
